Draft

### 2019-10-07 draft.tex
Modalities expanded

	For every $x, y : \| A\|$ we have $x = y$.

@@

 	We form a notion of \textbf{equivalence of types} based on maps with contractible fibers. 
 	Whenever we have some $f : A\rightarrow B$ with contractible fibers, we say that the types $A$ and $B$ are equivalent and write $A \simeq B$. 
 	To motivate this, note that we can form $g : B \rightarrow A$ so that $f\circ g \sim \mathtt{id}_B$ and $g\circ f \sim \mathtt{id}_A$. 
 	So we can use a witness of one of the types to get one of the other. 
 	Furthermore, the equivalence of types we introduced just now is an equivalence relation on $\mathcal{U}$. 
 	More exposition on equivalence can be found in chapter 4 of \cite{hottbook}.

 	We will now see some Type Theory variants that expand a bit upon what we've laid out in this section. 
 	Before embarking on that, we present the notion of \textbf{function extensionality} which we do not require to hold in the type theory presented in this chapter, yet holds in both of the variants below. 
 	As a prerequisite to it, we define the function 
 	$$\mathtt{happly} : (f = g) \rightarrow \prod_{a : A}(f(a) =_{B(a)} g(a))$$
 	for arbitrary functions $f,g : \prod_{a :A}B\;a$. We do so using induction on the identity path $f = g$ which reduces it to providing a witness for $f\;a =_{B(a)} g\;a$ with $f \equiv g$. 
 	Clearly $f\;a \equiv g\;a$ so $\mathtt{idp}_{f(a)} : f\;a =_{B(a)} g\;a$. 
 	\begin{axiom}[Function extensionality] 
 	$\mathtt{happly}$ has contractible fibers. 
 	\end{axiom}
 	Function extensionality, when true, allows us to equate functions that agree on all inputs.

@@

We would also like a way to do classical logic in Type Theory. 
To do so we would need a way to say that some type $C$ is inhabited without providing a specific witness $c : C$, which would have the undesirable effect of providing more information than a classical statement of existence. 
An effective way to do so is to provide a witness of the truncation $\lVert C \rVert$ and say that $C$ is \textit{merely} inhabited. 
So to work classically we restrict ourselves to using only mere propositions. 
This approach is called `propositions as mere propositions' in \cite{hottbook}. 
In the following, $A$ and $B$ are mere propositions.
\begin{itemize}
\item When we talk of conjunction $A \wedge B$, we mean the product $A \times B$. 
\item We interpret $A \vee B$, as the truncation $\| A + B\|$.
\item We interpret $\forall a\in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\prod_{a : A}P(a)$. 
\item We interpret $ \exists a \in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\| \sum_{a : A} P(a) \|$
\end{itemize}
Note that the above are all chosen so that they preserve mere propositions, e.g. $A\times B$ is a mere proposition. 
The use of the above notation is interchangeable in the sections to follow. 
We will try to specify whether a type is \textit{merely} or \textit{actually} inhabited avoid ambiguity whenever needed.

\chapter{Modalities}
%Under the view of types as $(\infty, 1)$-categories, functions between them are functors. 
%We can then view the universe $\mathcal{U}$ itslef as a category. 
\begin{defn}\label{modality_definition}
A \textbf{modality} is a function $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$ with the following properties.
\begin{enumerate}
	\item For every type $A$ we have a function $\eta_A^\bigcirc : A \rightarrow \bigcirc\; A$
	\item for every $A : \mathcal{U}$ and every type family $B : \bigcirc A \rightarrow \mathcal{U}$ we have a function
	$$\mathtt{ind}_\bigcirc : \Big( \prod_{a : A}\bigcirc (B (\eta_A^\bigcirc\; a)) \Big) \rightarrow \prod_{z : \bigcirc A} \bigcirc (B\; z)$$
	\item For every $f : \prod_{a : A} \bigcirc (B(\eta_A^\bigcirc\; a))$ and every $a : A$, there is a path $\mathtt{ind}_\bigcirc (f)(\eta_A^\bigcirc\; a) = f\; a$
	\item For all $z,z' : \bigcirc \; A$, the function $\eta_{z=z'}^\bigcirc : (z = z') \rightarrow \bigcirc (z = z')$ is an equivalence.

\end{enumerate}
\end{defn}
A modality $\bigcirc$ induces a $\Sigma$-closed reflective subuniverse.

\begin{defn}
Given modality $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$, a \textbf{reflective subuniverse} is a `subset' of $\mathcal{U}$ encoded by a family of h-propositions $P : \mathcal{U} \rightarrow \mathtt{Prop}$ such that the following conditions hold. 
\begin{itemize}
	\item For $A : \mathcal{U}$, we have $P(\bigcirc A)$.
	\item For $A : \mathcal{U}$ and $B$ such that $P(B)$, the function
	$$\lambda f.f \circ \eta_A^\bigcirc : (\bigcirc A \rightarrow B) \rightarrow (A \rightarrow B)$$
	is an equivalence.

\end{itemize}
The subuniverse is $\Sigma$\textbf{-closed} if for $X$ such that $P(X)$ and $Q : X \rightarrow \mathcal{U}$ such that $\prod_{x : X} P(Q(x))$, we have $P(\Sigma_{x : X} Q(x))$.
\end{defn}

\begin{thm}
Reflective subuniverses are closed under dependent products. 
That is for subuniverse $P$ and $B : A \rightarrow \mathcal{U}$ such that $\prod_{a : A}P(B(a))$, we have that $P(\prod_{a : A} B(a))$. 
\end{thm}

\begin{proof}
For $a : A$, consider $\mathtt{ev}_a: (\prod_{a: A}B(a)) \rightarrow B(a)$ defined by $\mathtt{ev}_a (f) : \equiv f(x)$. 
Since $P(B(a))$, we have 
$$(\lambda f.f \circ \eta_{\prod_{a: A}B(a)}^\bigcirc)^{-1} (\mathtt{ev}_a) : \bigcirc\big(\prod_{a: A}B(a)\big) \rightarrow B(a)$$
We can now define the retraction of $\eta^\bigcirc_{\prod_{a : A} B(a)}$ by pattern matching as such:\\
For $z: (\prod_{a: A}B(a))$ and $a : A$ we have 
$$(\lambda f.f \circ \eta_{\prod_{a: A}B(a)}^\bigcirc)^{-1} (\mathtt{ev}_a) (z) : B(a)$$

@@

\begin{lem}\label{405}
Under Church's thesis, the following type is inhabited
$$\sum_{F : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbf{2}} \Big( \big(\prod_{m : \mathbb{N}} \mathtt{atMost1one}\; F(m) \big) \times \big( \prod_{f : \mathbb{N} \rightarrow \mathbf{2}} \big\lVert \sum_{m, k : \mathbb{N}} F(m, 2k+ f\; m) = 1 \big\rVert \big) \Big)$$
\end{lem}
\begin{proof}
First we provide a witness
$$G : \sum_{F : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbf{2} } \prod_{n : \mathbb{N}}\mathtt{atMost1one}\; F(n)$$
For $n, m : \mathbb{N}$ we pick the indexed $T_n$ Turing machine of our enumeration. 
We can decide whether $m$ is odd or even. 
\begin{itemize}
\item If it's odd, then there exists actual $k : \mathbb{N}$ such that $m = 2*k +1$. 
If $k$ is the Gödel number of the computation history that $T_n$ goes through when given $n$ as the input, furthermore, if it halts at the end of this computation with output $1$ then set $G.\mathtt{fst}\; n\; m :\equiv 1$. 
Otherwise set $G.\mathtt{fst}\; n\; m :\equiv 0$.
\item If it's even, then there exists actual $k$ such that $m = 2*k$. 
We work as in the odd case, with the only difference being that we set $G.\mathtt{fst}\; n\; m :\equiv 1$ iff the output of the halting computation is $0$.
\end{itemize}
Having defined $G.\mathtt{fst}$ it's easy to see that $G.\mathtt{snd}$ has a straightforward proof, which we ommit.

@@

\chapter*{Appendix}
Following \cite{bridges_richman_1987_2} we develop a constructive theory of rational and real numbers.
\section{Rational numbers}
We define the rationals as $\mathbb{Q} :\equiv \sum_{n,m: \mathbb{N}}m \neq 0$ where $m$ is the denominator and $n$ the numerator. 
We now need to codify the usual notion of equality of rationals into type theory. 
We define it as the infix predicate
$$\doteq\; :\equiv \prod_{q_1, q_2 : \mathbb{Q}} (q_1.\mathtt{fst} \cdot q_2.\mathtt{snd}) = (q_2.\mathtt{fst} \cdot q_1.\mathtt{snd})$$
which can be proved to be an equivalence relation. 
This sort of equality between rationals is the only one that interests us and at no point will we be refering to the type theoretic primitive equality between $\mathbb{Q}$ witnesses, without explicit mention that we are doing so.

>>>>

	Finally, $\lVert A\rVert$ is a mere proposition.

@@

 	We form a notion of \textbf{equivalence of types} based on maps with contractible fibers. 
 	Whenever we have some $f : A\rightarrow B$ with contractible fibers, we say that the types $A$ and $B$ are equivalent and write $A \simeq B$ which we define as 
 	$$A\simeq B :\equiv \sum_{f : A \rightarrow B}\mathtt{hasContrFibers}\;f$$ 
 	To motivate this, note that whenever we have $A\simeq B$, we can form $f : A\rightarrow B$ and $g : B \rightarrow A$ so that $f\circ g \sim \mathtt{id}_B$ and $g\circ f \sim \mathtt{id}_A$. 
 	$f$ and $g$ are called each other's quasi-inverse.
 	It also holds the other way around, if there exist $f,g$ like above, then both of them have contractible fibers. 
 	Another way to show that $A\simeq B$ is to provide some $f : A\rightarrow B$ and $g_1,g_2 : B \rightarrow A$ such that $f \circ g_1 \sim \mathtt{id}_B$ and $g_2 \circ f \sim \mathtt{id}_A$. 
 	In these situations, $g_1$ is called a right inverse of $f$ and $g_2$ a left inverse. 
 	Finally, the equivalence of types we introduced just now is an equivalence relation on $\mathcal{U}$. 
 	More exposition on equivalence can be found in chapter 4 of \cite{hottbook}.

 	We will now see some Type Theory variants that expand a bit upon what we've laid out in this section. 
 	Before embarking on that, we present the notion of \textbf{function extensionality} which we do not require to hold in the type theory presented in this chapter, yet holds in both of the variants below. 
 	As a prerequisite to it, we define the function 
 	$$\mathtt{happly} : (f = g) \rightarrow (f \sim g)$$
 	for arbitrary functions $f,g : \prod_{a :A}B\;a$, where $A : \mathcal{U},\; B : A \rightarrow \mathcal{U}$. We do so using induction on the identity path $f = g$ which reduces it to providing a witness for $f\;a =_{B(a)} g\;a$ assuming $f \equiv g$. 
 	Clearly $f\;a \equiv g\;a$ so $\mathtt{idp}_{f(a)} : f\;a =_{B(a)} g\;a$. 
 	\begin{axiom}[Function extensionality] 
 	$\mathtt{happly}$ has contractible fibers. 
 	\end{axiom}
 	Function extensionality, when true, allows us to equate functions that agree on all inputs.

@@

We would also like a way to do classical logic in Type Theory. 
To do so we would need a way to say that some type $C$ is inhabited without providing a specific witness $c : C$, which would have the undesirable effect of providing more information than a classical statement of existence. 
An effective way to do so is to provide a witness of the truncation $\lVert C \rVert$ and say that $C$ is \textit{merely} inhabited. 
So to work classically we restrict ourselves to using only mere propositions. 
This approach is called `propositions as mere propositions' in \cite{hottbook}. 
\begin{itemize}
\item When we talk of conjunction $A \wedge B$, where $A$ and $B$ are mere propositions, we mean the product $A \times B$. 
\item We interpret $A \vee B$, where $A$ and $B$ are mere propositions, as the truncation $\| A + B\|$.
\item We interpret $\forall a\in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\prod_{a : A}P(a)$. 
\item We interpret $ \exists a \in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\| \sum_{a : A} P(a) \|$
\end{itemize}
Note that the above are all chosen so that they preserve mere propositions, e.g. $A\times B$ is a mere proposition. 
The use of the above notation is interchangeable in the sections to follow. 
We will try to specify whether a type is \textit{merely} or \textit{actually} inhabited to avoid ambiguity whenever needed.

\chapter{Modalities}
%Under the view of types as $(\infty, 1)$-categories, functions between them are functors. 
%We can then view the universe $\mathcal{U}$ itslef as a category. 
\begin{defn}\label{modality_definition}
A \textbf{modality} is any function $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$ with the following properties.
\begin{enumerate}
	\item For every type $A$ we have a function $\eta_A^\bigcirc : A \rightarrow \bigcirc A$ called the modal unit.
	\item for every $A : \mathcal{U}$ and every type family $B : \bigcirc A \rightarrow \mathcal{U}$ we have the induction principle
	$$\mathtt{ind}_\bigcirc : \Big( \prod_{a : A}\bigcirc (B (\eta_A^\bigcirc\; a)) \Big) \rightarrow \prod_{z : \bigcirc A} \bigcirc (B\; z)$$
	where $A$ and $B$ are implicit arguments of $\mathtt{ind}_\bigcirc$ and can be derived from context. 
	\item For every $f : \prod_{a : A} \bigcirc (B(\eta_A^\bigcirc\; a))$ and every $a : A$, there is a path $\mathtt{ind}_\bigcirc (f)(\eta_A^\bigcirc\; a) = f\; a$
	\item For all $z,z' : \bigcirc \; A$, the function $\eta_{z=z'}^\bigcirc : (z = z') \rightarrow \bigcirc (z = z')$ is an equivalence.
\end{enumerate}
\end{defn}

\begin{lem}\label{modalUnitLeftInverse}
Given $A : \mathcal{U}$, if $\eta^\bigcirc_A : A \rightarrow \bigcirc A$ has a left inverse, then $A \simeq \bigcirc A$. 
\end{lem}
\begin{proof}
Assume the hypotheses of the lemma. 
We need to show that $A \simeq \bigcirc A$. 
We already have a left inverse so producing a right inverse for $\eta^\bigcirc_A$ would be enough to show equivalence. 
Let $f : \bigcirc A \rightarrow A$ be the left inverse, i.e. $f \circ \eta^\bigcirc_A \sim \mathtt{id}_A$. 
We then have $\eta^\bigcirc_A \circ f \circ \eta^\bigcirc_A \sim \eta^\bigcirc_A \circ \mathtt{id}_A$ and 
$\eta^\bigcirc_A \circ f \circ \eta^\bigcirc_A \sim \mathtt{id}_{\bigcirc A} \circ \eta^\bigcirc_A$. 
This translates to
$$h : \prod_{a : A}\eta^\bigcirc_A \circ f \;(\eta^\bigcirc_A\; a) = \mathtt{id}_{\bigcirc A} \; (\eta^\bigcirc_A\; a)$$
We define the following function by assuming $a : A$ and applying the relevant modal unit to $h\;a$ 
$$h' : \prod_{a : A}\bigcirc\big(\eta^\bigcirc_A \circ f \;(\eta^\bigcirc_A\; a) = \mathtt{id}_{\bigcirc A} \; (\eta^\bigcirc_A\; a) \big)$$
We then use the induction principle of the modality to get
$$\mathtt{ind}_{\bigcirc}\; h' : \prod_{z : \bigcirc A}\bigcirc\big(\eta^\bigcirc_A \circ f \;(z) = \mathtt{id}_{\bigcirc A} \; (z) \big)$$
Then, by the equivalence mentioned in the fourth datum of \ref{modality_definition} we have a quasi-inverse $r$ for the modal unit $\eta_{\eta^\bigcirc_A \circ f \;(z) = \mathtt{id}_{\bigcirc A} \; (z)}$, 
which we use to construct
$$ \lambda\;(z : \bigcirc A).\; r \big((\mathtt{ind}_{\bigcirc}\; h')\;(z)\big) : \prod_{z : \bigcirc A}\eta^\bigcirc_A \circ f \;(z) = \mathtt{id}_{\bigcirc A} \; (z) $$
We've proven $\eta^\bigcirc_A \circ f \sim \mathtt{id}_{\bigcirc A}$ which means that $ f$ is the right inverse of $\eta_A^\bigcirc$. 
Since the modal unit has both a left and a right inverse, we can conclude that $A \simeq \bigcirc A$. 
\end{proof}
The usefulness of this lemma lies in that it makes it easier for us to provide a proof of equivalence between a type and its image under $\bigcirc$, when they are so. 
These types are of special interest to us because they form a $\sum$-closed reflective subuniverse of $\mathcal{U}$. 
We shall define a predicate to tell them apart. 
\begin{defn}
We define $\mathtt{isModal}_{\bigcirc} : \mathcal{U} \rightarrow \mathtt{Prop}$ as such
$$\mathtt{isModal}_{\bigcirc} \equiv \prod_{A : \mathcal{U}}(A \simeq \bigcirc A)$$
\end{defn}

\begin{defn}
Given modality $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$, the \textbf{reflective subuniverse} of $\mathcal{U}$ is encoded by the following type
$$\mathcal{U}^\bigcirc \equiv \sum_{A : \mathcal{U}}\mathtt{isModal}_\bigcirc (A)$$
% \begin{itemize}
% 	\item For $A : \mathcal{U}$, we have $\mathtt{isModal}_{\bigcirc}(\bigcirc A)$.
% 	\item For $A : \mathcal{U}$ and $B$ such that $\mathtt{isModal}_{\bigcirc}(B)$, the function
% 	$$\lambda (f : \bigcirc A \rightarrow B).(f \circ \eta_A^\bigcirc) : (\bigcirc A \rightarrow B) \rightarrow (A \rightarrow B)$$
% 	is an equivalence.

% \end{itemize}
The subuniverse is $\Sigma$\textbf{-closed} if for $X$ such that $\mathtt{isModal}_{\bigcirc}(X)$ and $Q : X \rightarrow \mathcal{U}$ such that $\prod_{x : X} \mathtt{isModal}_{\bigcirc}(Q(x))$, we have $\mathtt{isModal}_{\bigcirc}(\Sigma_{x : X} Q(x))$.
\end{defn}
The \textit{reflective} modifier refers to the fact that for each $f : A \rightarrow B$ there is a canonical way to construct its \textit{reflection}, $f' : \bigcirc A \rightarrow \bigcirc B$. 
We simply compose $f$ with the modal unit $\eta_B^\bigcirc$ and use the induction principle of the modality on the result.

Propositional truncation is a modality as it possesses the required data outlined in \ref{modality_definition}. 
Its reflective subuniverse is the universe of mere propositions.  
\begin{prop}
For all $A  :\mathcal{U}$, we have $\mathtt{isModal}_{\bigcirc}(\bigcirc A)$. 
\end{prop}
\begin{lem}
For $A, B : \mathcal{U}$ such that $\mathtt{isModal}\; B$ we have
$$(A\rightarrow B) \simeq (\bigcirc A \rightarrow B)$$
\end{lem}
\begin{proof}
Note that $(A\rightarrow B) \simeq (A\rightarrow \bigcirc B)$ and $(\bigcirc A \rightarrow B) \simeq (\bigcirc A \rightarrow \bigcirc B)$, since $B$ is modal. 
So we reduce our goal to 
$$(A\rightarrow \bigcirc B) \simeq (\bigcirc A \rightarrow \bigcirc B)$$
We need only provide a quasi-inverse for 
$$(-\circ \eta_A^{\bigcirc}) : (A\rightarrow \bigcirc B) \rightarrow (\bigcirc A \rightarrow \bigcirc B)$$
 to conclude the proof. 
We propose $\mathtt{ind}$ as the quasi-inverse. 
We first show $\big(\mathtt{ind}\circ (-\circ \eta_A^{\bigcirc})\big) \sim \mathtt{id}_{\bigcirc A \rightarrow \bigcirc B}$. 
Let $g : \bigcirc A \rightarrow \bigcirc B$. 
We need $\mathtt{ind}(g\circ \eta_A^\bigcirc ) = g$. 
By function extensionality it's enough to show that 
$$\prod_{x : \bigcirc A}\mathtt{ind}(g\circ \eta_A^\bigcirc )(x) = g(x)$$
So, for every $x$, we are trying to provide a witness for an identity path of terms in $\bigcirc B$. 
By the fourth datum of \ref{modality_definition} and the relevant induction principle with, with implicit arguments $A$ and $x\mapsto \mathtt{ind}(g\circ \eta_A^\bigcirc )(x) = g(x) : \bigcirc A \rightarrow \mathcal{U}$, we can reduce it to constructing a witness for
$$\prod_{a : A} \mathtt{ind}(g\circ \eta_A^\bigcirc )(\eta^\bigcirc_{A}\;a) = g(\eta^\bigcirc_{A}\;a)$$
A witness of this type is secured for us, once again by the definition of modalities, the third item with $g \circ \eta_A^\bigcirc $ as $f$.

The proof of $\big( (-\circ \eta_A^{\bigcirc})\circ \mathtt{ind}\big) \sim \mathtt{id}_{A \rightarrow \bigcirc B}$ 
is similar to the above, with some steps being even more immediate. 
\end{proof}
It is easy to generalize the above lemma to dependent functions.
\begin{cor}
For $A : \mathcal{U}$ and $B : \bigcirc A \rightarrow \mathcal{U}$ \\
such that $\prod_{x : \bigcirc A}\mathtt{isModal}_\bigcirc (B\;x)$ we have
$$\Big(\prod_{a : A}B(\eta_A^\bigcirc \; a)\Big) \simeq \Big( \prod_{x : \bigcirc A}B(x) \Big)$$
\end{cor}

\begin{thm}
Reflective subuniverses are closed under dependent products.
That is, for the subuniverse of $\bigcirc$ and $B : A \rightarrow \mathcal{U}$ such that\\
$\prod_{a : A}\mathtt{isModal}_{\bigcirc}(B(a))$, we have that $\mathtt{isModal}_{\bigcirc}(\prod_{a : A} B(a))$. 
\end{thm}

\begin{proof}
By \ref{modalUnitLeftInverse} it is enough to provide a left inverse for $\eta^\bigcirc_{\prod_{a : A} B(a)}$. 
For $a : A$, consider $\mathtt{ev}_a: (\prod_{a: A}B(a)) \rightarrow B(a)$ defined by $\mathtt{ev}_a (f) : \equiv f(a)$. 
Since $\mathtt{isModal}_{\bigcirc}(B(a))$, we have 
$$(\lambda f.f \circ \eta_{\prod_{a: A}B(a)}^\bigcirc)^{-1} (\mathtt{ev}_a) : \bigcirc\big(\prod_{a: A}B(a)\big) \rightarrow B(a)$$
We can now define the retraction of $\eta^\bigcirc_{\prod_{a : A} B(a)}$ by pattern matching as such:\\
For $z: (\prod_{a: A}B(a))$ and $a : A$ we have 
$$(\lambda f.f \circ \eta_{\prod_{a: A}B(a)}^\bigcirc)^{-1} (\mathtt{ev}_a) (z) : B(a)$$

@@

\begin{lem}\label{405}
Under Church's thesis, the following type is inhabited
$$\sum_{F : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbf{2}} \Big( \big(\prod_{m : \mathbb{N}} \mathtt{atMost1one}\; F(m) \big) \times \big( \prod_{f : \mathbb{N} \rightarrow \mathbf{2}} \big\lVert \sum_{m, k : \mathbb{N}} F(m, 2k+ f\; m) = 1 \big\rVert \big) \Big)$$
\end{lem}
\begin{proof}
First we provide a witness
$$G : \sum_{F : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbf{2} } \prod_{n : \mathbb{N}}\mathtt{atMost1one}\; F(n)$$
For $n, m : \mathbb{N}$ we pick the indexed $T_n$ Turing machine of our enumeration. 
We can decide whether $m$ is odd or even. 
\begin{itemize}
\item If it's odd, then there exists actual $k : \mathbb{N}$ such that $m = 2k +1$. 
If $k$ is the Gödel number of the computation history that $T_n$ goes through when given $n$ as the input, furthermore, if it halts at the end of this computation with output $1$ then set $G.\mathtt{fst}\; n\; m :\equiv 1$. 
Otherwise set $G.\mathtt{fst}\; n\; m :\equiv 0$.
\item If it's even, then there exists actual $k$ such that $m = 2k$. 
We work as in the odd case, with the only difference being that we set $G.\mathtt{fst}\; n\; m :\equiv 1$ iff the output of the halting computation is $0$.
\end{itemize}
Having defined $G.\mathtt{fst}$ it's easy to see that $G.\mathtt{snd}$ has a straightforward proof, which we ommit.

@@

\chapter*{Appendix}
Following \cite{bridges_richman_1987_2} we develop a constructive theory of rational and real numbers.
\section{Rational numbers}
We define the rationals as $\mathbb{Q} :\equiv \sum_{n,m: \mathbb{I}}m \neq 0$ where $m$ is the denominator and $n$ the numerator. 
We now need to codify the usual notion of equality of rationals into type theory. 
We define it as the infix predicate
$$\doteq\; :\equiv \prod_{q_1, q_2 : \mathbb{Q}} (q_1.\mathtt{fst} \cdot q_2.\mathtt{snd}) = (q_2.\mathtt{fst} \cdot q_1.\mathtt{snd})$$
which can be proved to be an equivalence relation. 
This sort of equality between rationals is the only one that interests us and at no point will we be refering to the type theoretic primitive equality between $\mathbb{Q}$ witnesses, without explicit mention that we are doing so.
